{ \sffamily \LARGE Electronic supplementary material}\\{\bf \sffamily \LARGE Evolution of reciprocity with limited payoff memory

### 2023-11-01 paper/main.tex

\begin{abstract}
\noindent
Direct reciprocity can explain the evolution of cooperative behaviors in repeated interactions. 
According to this literature, individuals should naturally learn to adopt conditionally cooperative strategies if they have multiple encounters with their interaction partner. 
Corresponding models have greatly facilitated our understanding of cooperation, yet they often make strong assumptions on how individuals remember and process payoff information. 
For example, when strategies are updated through social learning, it is commonly assumed that individuals compare their respective average payoffs.  
This would require them to compute (or remember) their payoffs against everyone else in the population.
To understand how more realistic constraints influence direct reciprocity, we consider the evolution of conditional behaviors when individuals learn based on more recent experiences.
Even in the most extreme case that they only take into account their very last interaction, we find that cooperation can still evolve. 
However, such individuals adopt less generous strategies, and they tend to cooperate less often than in the classical setup with expected payoffs. 
Interestingly, once individuals remember the payoffs of two or three recent interactions, cooperation rates quickly approach the classical limit. 
These findings contribute to a literature that explores which kind of cognitive capabilities are required for reciprocal cooperation. 
While our results suggest that some rudimentary form of payoff memory is necessary, it already suffices to remember a few interactions.
\end{abstract}

@@

In a one-shot donation game, players can only choose among two pure strategies (they can either cooperate or defect).
In the repeated game, strategies become arbitrarily complex. 
Here, strategies are contingent rules, telling players what to do depending on the outcome of all their previous interactions. 
For simplicity, in the following we assume individuals use {\it reactive strategies}~\citep{Nowak1992tit}. 
A reactive strategy only depends on the other player's action in the previous round. 
Such strategies can be written as a three-dimensional tuple \(\mathbf{s}=(y, p,q)\). 
The first entry \(y\) is the probability that the player opens with cooperation in the first round. 
The two other entries are the probabilities that the player cooperates in all subsequent rounds, depending on whether the co-player cooperated~($p$) or defected~($q$) in the previous round. 
The set of reactive strategies is simple enough to facilitate an explicit mathematical analysis~\citep{hofbauer1998evolutionary}. 
Yet it is rich enough to capture several important strategies of repeated games. 
For example, it contains \alld{} $=\!(0,0,0)$, the strategy that always defects. 
Similarly, it contains Tit-for-Tat, \tft{} $=\!(1,1,0)$, the strategy that copies the co-player's previous action (and that cooperates in the first round). 
Finally, it contains Generous Tit-for-Tat, \gtft $=\!(1,1,q)$, where $q\!>\!0$ reflects a player's generosity in response to a co-player's defection~\citep{molander:jcr:1985,Nowak1992tit}.

@@

We deviate from previous models in how we interpret the payoffs $\pi_\learner$ and $\pi_\rolemodel$, which form the basis of the pairwise comparisons in Eq.~\eqref{Eq:rho}. 
In previous work, these payoffs are taken to be the respective players' expected payoffs. 
We interpret that setup as a model with perfect payoff memory. 
There, the payoffs  $\pi_\learner$ and $\pi_\rolemodel$ represent an average over all possible repeated games the two individuals have played with all population members (\FigBaseResults, upper left panel). 
The use of expected payoffs is mathematically convenient, because explicit formulas for these payoffs are available~\citep{hofbauer1998evolutionary}.
Herein, we compare this model of perfect payoff memory to a model with limited payoff memory. 
In that model, the players' payoffs $\pi_\learner$ and $\pi_\rolemodel$ are taken to be the payoffs that each player received in their very last round prior to making social comparisons. 
That is, players only consider the very last repeated game they participated in, and there they only take into account the outcome of the very last round (\FigBaseResults, lower left panel). 
This assumption could reflect, for example, a strong recency bias in how individuals evaluate payoffs.  
In addition to this extreme case of limited payoff memory, later on we also explore cases in which players take into account the outcome of two, three, or four recent rounds.

@@

Both in the case of perfect and limited memory, we iterate the elementary strategy updating step described above for many time steps. 
This gives rise to a stochastic process that describes which strategies players adopt over time. 
We explore the dynamics of this process mathematically and with computer simulations.
For the results presented in the following, we assume that mutations are rare (\(\mu\!\rightarrow\! 0\)). 
This assumption is fairly common in evolutionary game theory, because it makes some computations more efficient~\citep{fudenberg:JET:2006,wu:JMB:2012,mcavoy:jet:2015}, and because the results can be interpreted more easily.  
However, in Section~9 of the \esm{} we show that our main results continue to hold for strictly positive mutation rates.

@@

\noindent 
{\bf The effect of increasing individual payoff memory.}
So far, we have taken a rather extreme interpretation of limited payoff memory. 
In the respective scenario, we assumed that individuals update their strategies based on their experience in a single round of the prisoner's dilemma, against a single co-player. 
The limited payoff memory framework can be expanded in various ways. 
In particular, individuals may recall a larger number of rounds, they may recall their interactions with several co-players, or both. 
To gain further insights on the impact of payoff memory, we explore three additional scenarios. 
In the first scenario, players recall the payoffs they obtained in the last two rounds against a single co-player. 
In the second scenario, players recall their last-round payoffs against two co-players. 
Finally, in the last scenario, they recall the two last rounds against two co-players (further extensions are straightforward, but we do not explore them here).

@@

For the first two scenarios, we can derive an analytical condition for when cooperation is stochastically stable. 
To this end, we again assume populations are large and that selection is strong. 
For simplicity, we also assume that the game continues almost certainly after each round (i.e., $\delta$ approaches one). 
The details of this analysis can be found in the \esm. 
Interestingly, we find in both scenarios that for $b\!>\!2c$, cooperation is stochastically stable when $q\!<\! \frac{\sqrt{2}}{2}\approx 0.707$. 
Comparing this condition with the more stringent condition in Eq.~\eqref{Eq:LimitedMemory} suggests that there are now more conditionally cooperative strategies that can sustain cooperation. 
Hence, cooperation should evolve more easily.

@@

We explore this prediction with additional simulations, again for a low and a high benefit $b$, 
see \FigHigherMemory.
We observe that a minimal increase in the players' payoff memory (compared to the baseline case with a single round recalled) can promote cooperation considerably. 
Specifically, in all three scenarios with extended memory, we now see similar cooperation rates, and they approach the rates observed under perfect memory. 
These results suggest that while it takes {\it some} payoff memory to sustain substantial cooperation rates, the requirements on memory seem to be rather modest. 
Already remembering a few interactions, either with the same co-player or across different co-players, may provide players with enough information to adopt reciprocal strategies.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\textwidth]{static/more_memory_summary_results.pdf}
  \caption{{\bf Average cooperation rates for different payoff memories.}
  We vary how much information individuals take into account when updating their strategies. 
  From left to right, we consider the following cases.
  ({\it i}) Updating occurs based on expected payoffs (perfect memory), 
  ({\it ii}) it occurs based on the last round of one interaction (limited memory), 
  ({\it iii}) based on the last round of two interactions,
  ({\it iv}) based on the last two rounds of one interaction,
  ({\it v}) based on the last two rounds of two interactions, and
  ({\it vi}) based on the expected payoff of one interaction.
  Again, simulations are run either for a comparably high benefit of cooperation~($b/c\!=\!10$), or for a low benefit~($b/c\!=\!3$). 
  We observe that perfect memory always yields the highest cooperation rate. 
  However, when individuals take into account at least two past interactions -- cases ({\it iii}) to ({\it vi}) -- evolving cooperation rates are close to this optimum. 
  Baseline parameters are the same as in \FigDependenceParameters.}
\label{fig:cooperation_rate_all_updating_payoffs}
\end{figure}

>>>>

\begin{abstract}
\noindent
Direct reciprocity can explain the evolution of cooperative behaviors in repeated interactions. 
According to this literature, individuals should naturally learn to adopt conditionally cooperative strategies if they have multiple encounters with their interaction partner. 
Corresponding models have greatly facilitated our understanding of cooperation, yet they often make strong assumptions on how individuals remember and process payoff information. 
For example, when strategies are updated through social learning, it is commonly assumed that individuals compare their respective average payoffs.  
This would require them to compute (or remember) their payoffs against everyone else in the population.
To understand how more realistic constraints influence direct reciprocity, we consider the evolution of conditional behaviors when individuals learn based on more recent experiences.
Even in the most extreme case that they only take into account their very last interaction, we find that cooperation can still evolve. 
However, such individuals adopt less generous strategies, and they tend to cooperate less often than in the classical setup with average payoffs. 
Interestingly, once individuals remember the payoffs of two or three recent interactions, cooperation rates quickly approach the classical limit. 
These findings contribute to a literature that explores which kind of cognitive capabilities are required for reciprocal cooperation. 
While our results suggest that some rudimentary form of payoff memory is necessary, it already suffices to remember a few interactions.
\end{abstract}

@@

In a one-shot donation game, players can only choose among two pure strategies (they can either cooperate or defect).
In the repeated game, strategies become arbitrarily complex. 
Here, strategies are contingent rules, telling players what to do depending on the outcome of all their previous interactions. 
For simplicity, in the following we assume individuals use {\it reactive strategies}~\citep{Nowak1992tit}. 
A reactive strategy only depends on the other player's action in the previous round. 
Such strategies can be written as a three-dimensional tuple \(\strategy=(y, p,q)\).
The first entry \(y\) is the probability that the player opens with cooperation in the first round. 
The two other entries are the probabilities that the player cooperates in all subsequent rounds, depending on whether the co-player cooperated~($p$) or defected~($q$) in the previous round. 
The set of reactive strategies is simple enough to facilitate an explicit mathematical analysis~\citep{hofbauer1998evolutionary}. 
Yet it is rich enough to capture several important strategies of repeated games. 
For example, it contains \alld{} $=\!(0,0,0)$, the strategy that always defects. 
Similarly, it contains Tit-for-Tat, \tft{} $=\!(1,1,0)$, the strategy that copies the co-player's previous action (and that cooperates in the first round). 
Finally, it contains Generous Tit-for-Tat, \gtft $=\!(1,1,q)$, where $q\!>\!0$ reflects a player's generosity in response to a co-player's defection~\citep{molander:jcr:1985,Nowak1992tit}.

@@

We deviate from previous models in how we interpret the payoffs $\pi_\learner$ and $\pi_\rolemodel$, which form the basis of the pairwise comparisons in Eq.~\eqref{Eq:rho}. 
In previous work, these payoffs are taken to be the respective players' expected payoffs. 
We interpret that setup as a model with perfect payoff memory. 
There, the payoffs  $\pi_\learner$ and $\pi_\rolemodel$ represent an average over all possible repeated games the two individuals have played with all population members (\FigBaseResults, upper left panel). 
The use of expected payoffs is mathematically convenient, because explicit formulas for these payoffs are available~\citep{hofbauer1998evolutionary}.
Herein, we compare this model of perfect payoff memory to a model with limited payoff memory. 
In that model, the players' payoffs $\pi_\learner$ and $\pi_\rolemodel$ are taken to be the payoffs that each player received in their very last round prior to making social comparisons. 
That is, players only consider the very last repeated game they participated in, and there they only take into account the outcome of the very last round (\FigBaseResults, lower left panel). 
This assumption could reflect, for example, a strong recency bias in how individuals evaluate payoffs.  
In addition to this extreme case of limited payoff memory, later on we also explore cases in which players take into account the outcome of two, three, four, or more recent rounds.

@@

Both in the case of perfect and limited memory, we iterate the elementary strategy updating step described above for many time steps. 
This gives rise to a stochastic process that describes which strategies players adopt over time. 
We explore the dynamics of this process mathematically and with computer simulations.
For the results presented in the following, we assume that mutations are rare (\(\mu\!\rightarrow\! 0\)). 
This assumption is fairly common in evolutionary game theory, because it makes some computations more efficient~\citep{fudenberg:JET:2006,wu:JMB:2012,mcavoy:jet:2015}, and because the results can be interpreted more easily.  
However, in Section~3 of the \esm{} we show that our main results continue to hold for strictly positive mutation rates.

@@

\noindent 
{\bf The effect of increasing individual payoff memory.}
So far, we have taken a rather extreme interpretation of limited payoff memory. 
In the respective scenario, we assumed that individuals update their strategies based on their experience in a single round of the prisoner's dilemma, against a single co-player. 
The limited payoff memory framework can be expanded in various ways. 
In particular, individuals may recall a larger number of rounds, they may recall their interactions with several co-players, or both. 
To gain further insights on the impact of payoff memory, we explore four additional scenarios. 
In the first scenario, players recall the payoffs they obtained in the last two rounds against a single co-player. 
In the second scenario, players recall their last-round payoffs against two co-players. 
In the third scenario, they recall the two last rounds against two co-players. 
Finally, in the last scenario, players update based on the average payoff they
receive over all rounds with a single co-player (further extensions are straightforward, but we do not explore them here).

@@

For most scenarios, we derive an analytical condition for when cooperation is stochastically stable. 
To this end, we again assume populations are large and that selection is strong. 
For simplicity, we also assume that the game continues almost certainly after each round (i.e., $\delta$ approaches one). 
The details of this analysis can be found in the \esm. 
In the first two scenarios, we interestingly find that for $b\!>\!2c$, cooperation is stochastically stable when $q\!<\! \frac{\sqrt{2}}{2}\approx 0.707$. 
Comparing this condition with the more stringent condition in Eq.~\eqref{Eq:LimitedMemory} suggests that there are now more conditionally cooperative strategies that can sustain cooperation. 
Hence, cooperation should evolve more easily.
In the last scenario, we find that cooperation is stochastically stable when \(q
< 1 - \frac{c}{b}\), which is the same condition as in
Eq.~\eqref{Eq:PerfectMemory}, even when only a single co-player is considered
instead of the whole population. In this case, the evolved cooperation
approaches the perfect payoff memory case the most.

@@

We explore these prediction with additional simulations, again for a low and a high benefit $b$, 
see \FigHigherMemory.
We observe that a minimal increase in the players' payoff memory (compared to the baseline case with a single round recalled) can promote cooperation considerably. 
Specifically, in all four scenarios with extended memory, we now see similar cooperation rates, and they approach the rates observed under perfect memory. 
These results suggest that while it takes {\it some} payoff memory to sustain substantial cooperation rates, the requirements on memory seem to be rather modest. 
Already remembering a few interactions, either with the same co-player or across different co-players, may provide players with enough information to adopt reciprocal strategies.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\textwidth]{static/more_memory_summary_results.pdf}
  \caption{{\bf Average cooperation rates for different payoff memories.}
  We vary how much information individuals take into account when updating their strategies. 
  From left to right, we consider the following cases.
  ({\it i}) Updating occurs based on expected payoffs (perfect memory), 
  ({\it ii}) it occurs based on the last round of one interaction (limited memory), 
  ({\it iii}) based on the last round of two interactions,
  ({\it iv}) based on the last two rounds of one interaction,
  ({\it v}) based on the last two rounds of two interactions, and
  ({\it vi}) based on the average payoff of one interaction.
  Again, simulations are run either for a comparably high benefit of cooperation~($b/c\!=\!10$), or for a low benefit~($b/c\!=\!3$). 
  We observe that perfect memory always yields the highest cooperation rate. 
  However, when individuals take into account at least two past interactions -- cases ({\it iii}) to ({\it vi}) -- evolving cooperation rates are close to this optimum. 
  Baseline parameters are the same as in \FigDependenceParameters.}
\label{fig:cooperation_rate_all_updating_payoffs}
\end{figure}

### 2023-11-01 paper/si.tex

\definecolor{darkblue}{rgb}{0,0,.8}
\definecolor{darkgreen}{rgb}{0,0.5,0.1}
\definecolor{lightapricot}{rgb}{0.99, 0.84, 0.69}
\newcommand{\matlabfunc}[1]{\textcolor{darkblue}{#1}}
\newcommand{\matlabcomment}[1]{\textcolor{darkgreen}{#1}}
\newcommand{\christian}[1]{\textcolor{blue}{\textbf{CH}: #1}}
\newcommand{\alex}[1]{\textcolor{red}{\textbf{AL}: #1}}

@@

In the following, we discuss our different memory settings in more detail. 
We discuss six cases explicitly. In these cases, updating occurs
({\it i})~based on average payoffs based on all interactions (perfect memory), 
({\it ii})~based on the last round of one interaction (limited memory), 
({\it iii})~based on the last round of two interactions,
({\it iv})~based on the last two rounds of one interaction,
({\it v})~based on the last two rounds of two interactions, and
({\it v})~based on the expected payoff of one interaction.
In each case, we consider the case that there are only two strategies present in the population (a resident and a mutant strategy). 
We first derive how likely it is that a learner (of any type) assigns a given payoff $\pi_\learner$ to itself, and a payoff of $\pi_\rolemodel$ to the role model. 
This allows us to derive explicit expressions for $\lambda^-_k/\lambda^+_k$, and hence for the mutant's fixation probability according to Eq.~\eqref{eq:fixation_probability}.
Based on these expressions we can characterize under which conditions cooperation is stochastically stable.

@@

\subsection{Updates based on expected payoff of one repeated game}\label{section:one_interaction_full_memory}

@@

\noindent
{\bf Computing the ratio $\lambda^-_k/\lambda^+_k$.}
We revisit the population setup, consisting of \(k\) mutants with the strategy
\(\strategy_\mutant = (y_\mutant, p_\mutant, q_\mutant)\) and \(N-k\) residents
with strategies \(\strategy_\resident = (y_\resident, p_\resident,
q_\resident)\). In each step of the evolutionary process, we select a learner and
a role model. The learner or the role model can be either a resident, and the
other player is a mutant. We consider that the learner and the
model update based on their most recent interaction. As mentioned previously,
two major cases must be considered:

\begin{enumerate}
  \item The learner and the role model have had their most recent interaction with each other.
  \item The learner's last interaction was not with the role model.
\end{enumerate}

@@

We assume that the players' payoff received from their last interaction is their
expected payoff given by Eq.~\eqref{eq:expected_payoff}. Thus, a resident
interacting with a resident receives the payoff $\pi(\strategy_\resident, \strategy_\resident)$,
and $\pi(\strategy_\resident, \strategy_\mutant)$ when interacting with a mutant. Similarly, a
mutant interacting with a resident receives the payoff $\pi(\strategy_\mutant,
\strategy_\resident)$ and $\pi(\strategy_\mutant, \strategy_\mutant)$ when interacting with a mutant.\\

@@

\noindent
{\bf Stochastic stability of cooperation.}
We once again calculate how easily a single \alld{} mutant can invade into a
resident population of \gtft{} players. When resident interacts with another
resident or with a mutant,
the expected payoffs based on Eq.~\eqref{eq:expected_payoff} of the interactions are,
\begin{equation*}
\begin{array}{ll}
  \pi (\strategy_\resident, \strategy_\resident) =  b - c,
  &\pi (\strategy_\resident, \strategy_\mutant) =  - c \left(\delta \left(q - 1\right) + 1\right).
\end{array}
\end{equation*}
The expected payoff of the mutant, are
\begin{align*}
\begin{array}{ll}
  \pi (\strategy_\mutant, \strategy_\resident) =  b \left(\delta \left(q - 1\right) + 1\right),
  &\pi (\strategy_\mutant, \strategy_\mutant) =  0.
\end{array}
\end{align*}
Based on these expected payoffs, we compute the ratio
\begin{equation}
  \frac{\lambda^{-}_1}{\lambda^{+}_1} =
  \frac{\frac{1}{N-1}\frac{1}{\left(1 + e^{\beta \left(b \left(\delta q - \delta\right) + c\right)}\right)}
  + \frac{N - 2}{N-1}\frac{1}{\left(1 + e^{- \beta \left(- b \left(\delta q - \delta + 1\right) + b - c\right)}\right)}}
  {\frac{1}{N-1}\frac{1}{\left(1 + e^{- \beta \left(b \left(\delta q - \delta\right) + c\right)}\right)}
  + \frac{N - 2}{N-1}\frac{1}{\left(1 + e^{- \beta \left(b \left(\delta q - \delta + 1\right) - b + c\right)}\right)}}.
\end{equation}

@@

\christian{Adjust symbols in main text -- for role model, strategy, etc}
\end{document}

>>>>

\definecolor{darkblue}{rgb}{0,0,.8}
\definecolor{darkgreen}{rgb}{0,0.5,0.1}
\definecolor{lightapricot}{rgb}{0.99, 0.84, 0.69}
\newcommand{\matlabfunc}[1]{\textcolor{darkblue}{#1}}
\newcommand{\matlabcomment}[1]{\textcolor{darkgreen}{#1}}
\newcommand{\christian}[1]{\textcolor{blue}{\textbf{CH}: #1}}
\newcommand{\alex}[1]{\textcolor{red}{\textbf{AL}: #1}}
\newcommand{\nikoleta}[1]{\textcolor{orange}{\textbf{NG}: #1}}

@@

In the following, we discuss our different memory settings in more detail. 
We discuss six cases explicitly. In these cases, updating occurs
({\it i})~based on average payoffs based on all interactions (perfect memory), 
({\it ii})~based on the last round of one interaction (limited memory), 
({\it iii})~based on the last round of two interactions,
({\it iv})~based on the last two rounds of one interaction,
({\it v})~based on the last two rounds of two interactions, and
({\it v})~based on the average payoff of one interaction.
In each case, we consider the case that there are only two strategies present in the population (a resident and a mutant strategy). 
We first derive how likely it is that a learner (of any type) assigns a given payoff $\pi_\learner$ to itself, and a payoff of $\pi_\rolemodel$ to the role model. 
This allows us to derive explicit expressions for $\lambda^-_k/\lambda^+_k$, and hence for the mutant's fixation probability according to Eq.~\eqref{eq:fixation_probability}.
Based on these expressions we can characterize under which conditions cooperation is stochastically stable.

@@

\subsection{Updates based on average payoff of one repeated game}\label{section:one_interaction_full_memory}

@@

\nikoleta{CH to write a few senteces}

\noindent
{\bf Computing the ratio $\lambda^-_k/\lambda^+_k$.}
We revisit the population setup, consisting of \(k\) mutants with the strategy
\(\strategy_\mutant = (y_\mutant, p_\mutant, q_\mutant)\) and \(N-k\) residents
with strategies \(\strategy_\resident = (y_\resident, p_\resident,
q_\resident)\). In each step of the evolutionary process, we select a learner and
a role model. The learner or the role model can be either a resident, and the
other player is a mutant. We consider that the learner and the
model update based on a single repeated game with a member of the population.
As mentioned previously, two major cases must be considered:

\begin{enumerate}
  \item The learner and the role model play a repeated game with each other.
  \item The learner's last repeated game was not with the role model.
\end{enumerate}

@@

We assume that the players' calculations are based on the average payoffs (as given by
Eq.~\eqref{eq:expected_payoff})
received from interacting with their last interaction partners. Thus, a resident
interacting with a resident receives the payoff $\pi(\strategy_\resident, \strategy_\resident)$,
and $\pi(\strategy_\resident, \strategy_\mutant)$ when interacting with a mutant. Similarly, a
mutant interacting with a resident receives the payoff $\pi(\strategy_\mutant,
\strategy_\resident)$ and $\pi(\strategy_\mutant, \strategy_\mutant)$ when interacting with a mutant.\\

@@

\noindent
{\bf Stochastic stability of cooperation.}
We once again calculate how easily a single \alld{} mutant can invade into a
resident population of \gtft{} players. When a resident interacts with another
resident or with a mutant,
the average payoffs based on Eq.~\eqref{eq:expected_payoff} of the interactions are,
\begin{equation*}
\begin{array}{ll}
  \pi (\strategy_\resident, \strategy_\resident) =  b - c,
  &\pi (\strategy_\resident, \strategy_\mutant) =  - c \left(\delta \left(q - 1\right) + 1\right).
\end{array}
\end{equation*}
The average payoffs of the mutant, are
\begin{align*}
\begin{array}{ll}
  \pi (\strategy_\mutant, \strategy_\resident) =  b \left(\delta \left(q - 1\right) + 1\right),
  &\pi (\strategy_\mutant, \strategy_\mutant) =  0.
\end{array}
\end{align*}
Based on these average payoffs, we compute the ratio
\begin{equation}
  \frac{\lambda^{-}_1}{\lambda^{+}_1} =
  \frac{\frac{1}{N-1}\frac{1}{\left(1 + e^{\beta \left(b \left(\delta q - \delta\right) + c\right)}\right)}
  + \frac{N - 2}{N-1}\frac{1}{\left(1 + e^{- \beta \left(- b \left(\delta q - \delta + 1\right) + b - c\right)}\right)}}
  {\frac{1}{N-1}\frac{1}{\left(1 + e^{- \beta \left(b \left(\delta q - \delta\right) + c\right)}\right)}
  + \frac{N - 2}{N-1}\frac{1}{\left(1 + e^{- \beta \left(b \left(\delta q - \delta + 1\right) - b + c\right)}\right)}}.
\end{equation}

@@

\end{document}
